Extended Abstract

### 2021-05-10 paper/paper.tex
A little on inference

\newcommand{\picalc}{$\pi$-calculus}
\newcommand{\Picalc}{$\pi$-Calculus}
\newcommand{\rulename}[1]{{\tiny \textsc{(#1)}}}
\newcommand{\constr}[1]{\textcolor{olive}{\mathtt{#1}}}
\newcommand{\type}[1]{\textcolor{blue}{\mathtt{#1}}}

% Syntax types
\newcommand{\Fin}[1]{\type{Fin}~#1}
\newcommand{\fzero}{\constr{zero}}
\newcommand{\fsuc}{\constr{suc}}
\newcommand{\sExpr}[1]{\type{Expr}~#1}
\newcommand{\sProc}[1]{\type{Proc}~#1}
\newcommand{\tvar}[2]{#1 ~\type{\ni_t}~ #2}
\newcommand{\tkind}[2]{#1 ~\type{\vdash_t}~ #2}
\newcommand{\ttype}[1]{\type{Type}~#1}
\newcommand{\tusage}[1]{\type{Usage}~#1}
\newcommand{\tCtx}[2]{\type{Ctx}_{#1}~#2}
\newcommand{\tSplit}[3]{#1~\type{=}~#2~\type{\uplus}~#3}
\newcommand{\tun}[1]{\type{un}~#1}
\newcommand{\tVar}[3]{#1 ~ \type{\ni} ~ #2 ~ \type{:} ~ #3}
\newcommand{\tExpr}[3]{#1 ~ \type{\vdash} ~ #2 ~ \type{:} ~ #3}
\newcommand{\tProc}[2]{#1 ~ \type{\vdash} ~ #2}

@@

@@

@@

Constraints take two forms: the binary $\eqconstr{S}{T}$, where $S$ and $T$ are meant to be unified, and the ternary $\sumconstr{S}{T}{R}$, where the usage annotations in $T$ and $R$ are meant to be added up to $S$.
Both these constraints are defined on usage annotations, types, and typing contexts.
Constraints of the former form can \emph{always} be eagerly solved, while constraints of the latter \emph{sometimes} cannot, and must be deferred.

\paragraph{Constraint Satisfaction}

The \([\![\_]\!]\) function interprets constraints into their type
counterparts \(S \equiv T\) and \(S = T + R\). Substituting into a
constraint substitutes pointwise.

@@

\[
\texttt{infer} : \forall (p : \texttt{Proc}~n). ~ \exists \gamma. ~ \texttt{Ctx}~n~\gamma \times \texttt{List}~(\texttt{Constr}~\gamma)
\]

>>>>

\newcommand{\picalc}{$\pi$-calculus}
\newcommand{\Picalc}{$\pi$-Calculus}
\newcommand{\rulename}[1]{{\tiny \textsc{(#1)}}}
\newcommand{\constr}[1]{\textcolor{olive}{\mathtt{#1}}}
\newcommand{\func}[1]{\textcolor{gray}{\mathtt{#1}}}
\newcommand{\type}[1]{\textcolor{blue}{\mathtt{#1}}}

% Syntax types
\newcommand{\Fin}[1]{\type{Fin}~#1}
\newcommand{\fzero}{\constr{zero}}
\newcommand{\fsuc}{\constr{suc}}
\newcommand{\sExpr}[1]{\type{Expr}~#1}
\newcommand{\sProc}[1]{\type{Proc}~#1}
\newcommand{\tvar}[2]{#1 ~\type{\ni_t}~ #2}
\newcommand{\tkind}[2]{#1 ~\type{\vdash_t}~ #2}
\newcommand{\ttype}[1]{\type{Type}~#1}
\newcommand{\tusage}[1]{\type{Usage}~#1}
\newcommand{\tCtx}[2]{\type{Ctx}_{#1}~#2}
\newcommand{\tSplit}[3]{#1~\type{=}~#2~\type{\uplus}~#3}
\newcommand{\tEq}[2]{#1~\type{\equiv}~#2}
\newcommand{\tun}[1]{\type{un}~#1}
\newcommand{\tVar}[3]{#1 ~ \type{\ni} ~ #2 ~ \type{:} ~ #3}
\newcommand{\tExpr}[3]{#1 ~ \type{\vdash} ~ #2 ~ \type{:} ~ #3}
\newcommand{\tProc}[2]{#1 ~ \type{\vdash} ~ #2}
\newcommand{\tConstr}[1]{\type{Constr} ~ #1}
\newcommand{\tConstrs}[1]{\type{[Constr} ~ #1 ~ \type{]}}

@@

\newcommand{\subst}[2]{#1 ~ \func{\triangleleft} ~ #2}

@@

(Variables are black, $\type{\textrm{types are blue}}$, $\constr{\textrm{constructors are olive green}}$, $\func{\textrm{functions are gray}}$, and I am sorry this is not a wee poem.)

@@

\paragraph{Constraints}

Constraints are of type $\tConstr{\gamma}$ and are defined on arguments of type $\tkind{\gamma}{k}$ for some $k$ --- that is, both usage annotations and types.
They take two forms: the binary $\eqconstr{S}{T}$, where $S$ and $T$ are meant to be unified, and the ternary $\sumconstr{S}{T}{R}$, where $T$ and $R$ are meant to be added up to $S$.
Constraints of the former form can \emph{always} be eagerly solved, while constraints of the latter \emph{sometimes} cannot, and must be deferred.
We use $\tConstrs{\gamma}$ to refer to lists of constraints of type $\tConstr{\gamma}$.

\paragraph{Substitution}

We encode substitution as a function $\subst{\_}{\_}$ of type $(\forall k \to \tvar{\gamma}{k} \to \tkind{\delta}{k}) \to (\forall k \to \tkind{\gamma}{k} \to \tkind{\delta}{k})$ that acts on usage annotations and types and typing contexts, and that freely alters the metavariable context.
We define substitutions on constraints pointwise on their arguments.

\paragraph{Constraint satisfaction}

We use a $\func{[\![\_]\!]}$ function to interpret constraints $\eqconstr{S}{T}$ and $\sumconstr{S}{T}{R}$ into their proof counterparts $\tEq{S}{T}$ and $\tSplit{S}{T}{R}$.

@@

We encode type inference as a function taking processes with $n$ free variables and returning a metavariable context $\gamma$, a typing context with $n$ free variables containing metavariables in $\gamma$, and a list of constraints on those metavariables $\gamma$:
$$
\func{infer} : \sProc{n} \to ~ \exists \gamma. ~ \tCtx{n}{\gamma} \times \tConstrs{\gamma}
$$

\todo{insert examples}

We define a similar function for type inference on expressions, this time also returning a type $\tau$ of type $\ttype{\gamma}$ for the expression.

\todo{insert examples}
